Human Activity Recognition
========================================================

Objective of this analysis is to identify predictor variables and generate predictor algorithm to predict human activities performed based on realistic parameters.

Data used for this analysis is from Groupware@LES group. Please refer to their website for more details: http://groupware.les.inf.puc-rio.br/har

Let's start with loading libraries required for analysis

```{r}
library(RCurl)
library(caret)
```

Next, we need to load training and test data from sources

```{r}
trainurl<-getURL('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv', ssl.verifypeer = FALSE)
traindata<-read.csv(text=trainurl)
testurl<-getURL('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv', ssl.verifypeer = FALSE)
testdata<-read.csv(text=testurl)
```

Let's look at small subset of training to data to identify right predictors

```{r}
head(traindata)
table(traindata$classe)
```

We can remove initial columns like username, time stamp etc as they cannot be used for prediction. Also we can remove columns with too many NAs.

```{r}
traindata<-traindata[,-(1:7)]
traindata<-subset(traindata, select=-which(colSums(is.na(traindata))>19215))
```

We still have many columns (variables) those cannot be used for prediction because of too many missing values. Let's keep only columns those can be used for prediction.

```{r}
traindata<-subset(traindata,select=c(1,2,3,4,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,42,43,44,54,55,56,57,58,59,60,61,62,63,64,65,66,76,77,78,79,80,81,82,83,84,85,86))
```

After some trial I have decided to use random forest algorithm to build prediction model. I have a dual core machine to run the algorithm so we will run it in parallel mode. But still considering large amount of data, it will take some time to build prediction model.
```{r}
predictor<-train(classe~.,data=traindata,method='rf', tcControl=trainControl(method = "cv", number = 4),allowParallel=TRUE)
```

Let's examine predictor model for accuracy and variance
```{r}
predictor
predictor$finalModel
```

Now let's remove unnecessary columns from test data so that we can run prediction model we just built.
```{r}
testdata<-testdata[,-(1:7)]
testdata<-subset(testdata, select=-which(colSums(is.na(testdata))>19))
```

Now we can run prediction model on test data to get predictions
```{r}
testdata$prediction<-predict(predictor,newdata=testdata)
testdata$prediction
```

